Sketch Recognition and Understanding

### 2009-08-16 trunk/draft/System.tex
after most changes

\section{Preprocessing}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\subsection{Choosing the Best Fit}
\label{sec:bestFit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. The system choose the minimum segmentation as the final segmentation of this stroke. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.
\subsection{Structural and geometrical Features(FS1)}
Features defines the structure of the geometrical symbol.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

@@

 The vector $\omega$ determines th orientation of a seperation plane. The scalar $b$ determines the offeset of the plane from the orgin. Lets assume that the datatpointa re linearly seperable, i.e. there are infinitely many possible seperating planes that correctly classify the training data. Figure\ref{fig:svm1} illustrates two different seperating planes. It is intuitively noted from the figure that the solid line is more likely to generalize more data and thus introduce less classification error. This can be geometrically characterized as plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.

The SVM problem is how to construct the plane that is furthuest from both classes. One approcah is to maximize the margin between two parallel planes supporting. A plane supports a class if all points in that class are on one side of that plane. For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label. Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{fig:svm2}. The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class. The support vectors in Figure \ref{fig:svm2} are filled datapoints , 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

The distance or margin between these supporting planes 
\begin{equation}
 \omega \cdot x=b+1 \\
  \omega \cdot x=b-1
\end{equation}
is 
\begin{equation}
 \gamma = \frac{2}{\parallel \omega \parallel^2}
\end{equation}
Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ int the following quadratic program

@@

The margin depends on the Support vecotrs that are bumpt into the planes.  This is a primal problem Quadratic Programming (QP) problem. The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}
Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

\subsubsection{Theoretcical foundation}
From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained. Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function . For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity. Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability. The size of the margin is not directly dependent on the dimensionaliuty of the data. Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute). In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced. The problem is refered to the large volume of literature in, e.g., \cite{bookSVMoverfit11,statisticalSVM1} for more technical discussion of the taht statistical theroy.

We can gain inishgit into thesse result susing geomertic arguments. Classification function hat have capcity to fit the training data are more likely to overfit  resulting in poor generalization.

\subsubsection{Linearly Inseperable Case}
 So far we have assumed that the two datasets are linearly separable. This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail. As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails. Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains. Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Any point fallihng on the wrong side of tis supprting plane is considered to be an error. We want to maximize the margin and minimize the error.

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseperable}
	\label{fig:svm5}
\end{figure}

 This could be accomplished through minor change in the supporging plane QP problem. A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

Once again the primal relaxed supprting plane method is equivilent to the dual problem. The Largangian dual of the QP task is :

@@

\subsubsection {Non linear function Using Kernels}
If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{fig:svm6} ). A quadratic or  cubic function will work better in some casses. A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data. Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space. To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially. The first problem is that overfitting becomes a problem. SVM are largely immune to this porlblem since they rely on margin maximization. Provided an appropriate value of paaramter $C$ is chosen. The second concern is that it is not practical to actually compute $\vartheta(x)$. SVMs get around this issue through teh use of kernels.

>>>>

\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The preprocessing step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\subsection{Choosing the Best Fit}
\label{sec:bestFit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. The system choose the minimum segmentation as the final segmentation of this stroke.

After the best fit segmentaion is choosen the paramter of each primitive is computed. For example, if the stroke is represented as an ellispe, the center and axii of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and paramter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

After the segmentation, each stroke is represented as a list of segments $L_s$. A segment can be either an ellipse, a line or a circular arc. The list of segments $L_s$ represents the segmentaion of the storke. When a user draw another stroke the generated segmentaion of the storke is appended to the list of segments.  When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol. The relation and properties between different segments are extracted from the segment list $L_s$.  This step is considered as feature extration which is dircribed in details in the next sections.

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\subsection{Structural and geometrical Features(FS1)}
Features defines the structure of the geometrical symbol.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

@@

 The orientation of the speration plane is deteremined using the vector $\omega$ and the offest of the plane from the orgin is determined using the scalar $b$.   Lets make an assumption that the datatpoint are linearly seperable, i.e. there are infinite possibilities to find a plane that seperates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different seperating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.  plane supports a class if all points in that class are on one side of that plane.

The main problem of SVM is constructing the plane that is furthuest from both classes. The main approach is to use two parallel supporting planes and maximize the margin between them. All point in a class should be in one side of the plane in order to be a supporting plane.  Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. Similary the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}. To accomplish this, both planes are pushed apart untill they bump into small number of datapoints ( the support vecotrs ) for each class. Figure \ref{fig:svm2} shows the support vectors as filled datapoints, 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

The distance between these supporting planes 
\begin{equation}
 \omega \cdot x=b+1 \\
  \omega \cdot x=b-1
\end{equation}
is 
\begin{equation}
 \gamma = \frac{2}{\parallel \omega \parallel^2}
\end{equation}

Thus maximizing the margin is equivalent to minimizing $\frac{{\parallel \omega \parallel}^2}{2}$ in the next quadratic program

@@

The Support vectors that are pumped into the supprting plane deteremine the margin. This is a primal form Quadratic Programming (QP) problem. The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}

Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

\subsubsection{Theoretcical foundation}

The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.  The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexiy or capacity of the classification. In linear functions, using planes to maximize the margin of seperation reduces the complexiy of the functions. Thus, by explicity maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability. The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute). In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.  A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theroy.

%Using geomteric arguments we can gain in sights to some of these results. Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.

\subsubsection{Linearly Inseperable Case}
In all previous sections we assumed that the problem of classifiying two set is linearly separable. The assumption is not true in most casses, thus, trying to construct a plane to seperate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can seperate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseperable case.  Consider the linear inseperable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point missclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseperable}
	\label{fig:svm5}
\end{figure}
A minor change in the supporting plane QP problem can accomplish this. A nonegative slack or error variable $z_i$ is added to each constraint.  A weighted penalty term is also added to the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

The primal relaxed supporting plane method is equivalent to the dual problem. The Largangian dual of the QP task is :

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be seperated using linear space. A higher order plane can work better in similar casses. Converting a linear classification algorithm into non linear classification algoirhtm  can be achieved by adding addional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be seperated using an  existing linear classification algorithm.  The  linear classication that is applied on the feautre space produce a non linear classification on the orginal inputspace. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimentional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially. The first problem is that overfitting becomes a problem. However, since SVM rely on margin maximization they are largely immune to this porblem. Provided an appropriate value of paramter $C$ is chosen. The second concern is that it is the computaional complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.
